Ecosystem carbon balance in the Hawaiian Islands under different scenarios of future climate and land use change

### 2020-08-05 Manuscript/HI_Model_ms.Rmd
updated manuscript files

## Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under four scenarios of climate and land-use change. The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability. Simulations were run for 90 years at an annual timestep, using initial conditions corresponding to the year 2010. All simulations were repeated for 30 Monte Carlo realizations.

@@

 Transition targets were based on historical trends of land use change from 1992-2011 [@noaa_coastal_2020] and from population projections [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion of *Eucalyptus* spp. plantations to agriculture. The remaining 40% were rotation harvests.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016. Historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1) were based on a new 21-year historical wildfire spatial database of the Hawaiian Islands covering the years 1999 to 2019 (Figure S2). The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the 21-year median area burned statewide; the high land-use scenario sampled from historical fire years above this 21-year median (Fig. S2a).

>>>>

The main Hawaiian Islands are a complex mosaic of natural and human-dominated landscapes,

## Methods
We used the Land Use and Carbon Scenario Simulator (LUCAS), an integrated landscape change and carbon gain-loss model, to project changes in ecosystem carbon balance for the seven main Hawaiian Islands under all combinations of two land-use scenarios (low and high) and two radiative forcing scenarios (RCP 4.5 and RCP 8.5). The landscape change portion of LUCAS is a state-and-transition model that applies a Monte Carlo approach to track the state type and age of each simulation cell in response to a pre-determined set of transitions [@daniel_state-and-transition_2016]. The carbon gain-loss portion tracks carbon stocks within each simulation cell over time as continuous state variables, along with a pre-defined set of continuous flows specifying stock level rates of change over time [@daniel_integrating_2018; @sleeter_effects_2019]. We parameterized the Hawai`i LUCAS model to estimate annual changes in carbon stocks and fluxes in response to land use, land use change, wildland fire, and long-term climate variability. Simulations were run for 90 years at an annual timestep, using initial conditions corresponding to the year 2010. All simulations were repeated for 30 Monte Carlo realizations.

@@

 Transition targets were based on historical trends of land use change in the Hawaiian Islands from 1992-2011 [@noaa_coastal_2020] and on population projections for the State of Hawaii [@kim_population_2018]. For the high land-use scenario, transition rates for each timestep and Monte Carlo realization were sampled from uniform distributions bounded by the median and maximum historical rates of agricultural contraction, agricultural expansion, and urbanization for each island. For the low land-use scenario, rates of agricultural contraction and expansion were sampled from uniform distributions bounded by zero and the minimum historical rates for each island. Urbanization rates in the low land-use scenario were based on island-level population estimates and projections at five year intervals from 2010-2045 [@kim_population_2018]. We converted population projections into urbanization transition targets following @sleeter_future_2017 by calculating population density for each island and then projecting future developed area based on the five-year incremental change in island population. The spatial extent of agricultural contraction, agricultural expansion, and urbanization was constrained in both land-use scenarios based on existing zoning maps [@daniel_state-and-transition_2016]. Transition targets for tree plantation harvest were set at ~75% of recent historical rates in the high land-use scenario and ~40% of recent historical rates in the low land-use scenario [@daniel_state-and-transition_2016]. In both land-use scenarios, approximately 60% of tree plantation harvests were replacement harvests resulting in conversion of *Eucalyptus* spp. plantations to agriculture. The remaining 40% were rotation harvests with replanting of *Eucalyptus* spp.

The wildfire transition sub-model was modified from @daniel_state-and-transition_2016 by incorporating a new 21-year historical wildfire spatial database of the Hawaiian Islands (Figure S2) to calculate historical wildfire size distribution and ignition probabilities for each unique combination of moisture zone (Figure S1), island, and state type (Figure 1). The number and size of fires was randomly drawn from one of these historical year-sets for each timestep and Monte Carlo realization, using burn severity probabilities from @selmants_baseline_2017. Wildfire in the low land-use scenario was sampled from the subset of historical fire years at or below the median area burned statewide from 1999-2019, while the high land-use scenario sampled from historical fire years above the median area burned over the same 21-year period (Fig. S2a).
